Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-01 paper/paper.tex
updated paper with new outline

\title{Tweet Classification - Filtering \\ of Twitter for Company-relevant Tweets}
\author{
Curtis Ullerich, Daniel Stiner, Brandon Maxwell\\
Department of Computer Science and Engineering\\
Iowa State University
Ames, Iowa, USA\\
\{curtisu,stiner,bmaxwell\}@iastate.edu\\
}
\maketitle
\begin{abstract}
\begin{quote}
Twitter is a popular source for data mining due to its massive scale and inclusivity of current trends. As businesses seek to discover public opinion about their business or products, Twitter is one source for mining this information. Tweets present interesting classification challenges due to their irregular formatting and relatively small number of features. Tweet selection by keyword filtering often suffers from an high accuracy and low recall, or low accuracy and high recall. Using machine learning for text classification can improve this. We aim to survey a variety of text processing steps, focusing on the effects of tokenization, to increase accuracy of classification using a Naive Bayes classifier. We use a broad set of keywords with the goal of collecting a superset of the Tweets about a particular company (Apple, in this case) and train a binary classifier to filter out false positives.
\end{quote}
\end{abstract}

\section{Introduction}

\subsection{Goals}

\begin{itemize}

\item Collect a corpus that represents a superset of tweets about Apple and label these instances as "apple" or "none"
\item Create preprocessing steps to maximize the accuracy of binary classification

\end{itemize}

\subsection{Hypotheses}
\item Using URL-replacement will improve disambiguation accuracy. 
\item adding features for emoticon classes....
\item adding features for @references, retweets, and hashtags, will decrease the feature set while improving accuracy
\item selection of a proper tokenization regex will have profound effects on accuracy
\item Stopping and stemming....

\section{Related Work}
DUALIST is a system that extends Mallet by user-interactive feedback during training[Settles]. Included in their implementation was a TwitterPipe that performed some simple feature extraction. 
Part of the Sentiment Analysis Symposium focuses on tokenization of tweets for sentiment analysis, demonstrating that it provides consistent improvement over whitespace-based tokenizing \cite{potts2011}. A good deal of research has been done into sentiment classification of tweets \cite{Pak10}

\section{System Breakdown}
\textbf{Corpus creation}\\
By accessing the Twitter "Firehose" stream through their public API, we harvested tens of thousands of tweets using broad keyword filters with the goal of collecting a superset of relevant tweets. This provides us a smaller set of features in the overal corpus. The keywords used for the company of choice during this round of testing, Apple, were collected from information on the company's website and the Wikipedia page about Apple:\\
...\\
\\
This set of keywords clearly allows in a large number of false positives:\\
nice fish and chip.. good pie apple :-) http://t.co/KVDN5r5l \\
20 chicken nuggets, chips, big mac, and a vanilla milkshake LIFE IS GOOD \\
1 more day whoop plus safari ride yay \\
shley staples an apple.  "AHHHH!!!!  Apple juice in my eye!" \\
I ain't seen uncle Mac in a while\\
\\
During the time period of our data collection (give time period), the ratio was 40\% topical and 60\% false positives [need to analyze this for sure].
On Twitter, a single message posted is often reposted, or "re-tweeted," many more times in a very short timespan. To reduce the effects of temporal overfitting, in which our model would overfit the data for a particular time slice, we filtered the data to remove retweets and near-duplicates. Tweets automatically osted by applications, iPad games, for instance, are often near-duplicates, so we removed these from the data set as well. This reduced a collection of 100,000 tweets by 50\%. \\
A computer-posted tweet: I've collected 10,650 gold coins! http://t.co/H0V0O6yP #ipad, #ipadgames, #gameinsight\\
A retweet, signified by "RT": RT @PHILerNotebook: Trying to fix my Mac. This gray loading bar takes 10 minutes to start up. Grr. Shall take it to an _Apple_ _store_ soon! \\
\\
We hand labeled 2000 tweets for use in training and testing, including only English-language tweets in our training set. \\
Data sets and scripts used for processing text data are available on the project website.

\textbf{Data aggregation and classification}\\

\subsection{Basic Approach}

We do all machine learning with the Mallet[2] library developed by the University of Massachusetts. Mallet includes abstractions for a data processing pipeline, during which we use pre-built and custom processing "Pipes" to transform the data, beginning in our case with raw tweets text and ending in Mallet's internal FeatureVector format. We built several Pipes to analyze the effect of different filters and processes on the accuracy of classification:
SpellCheck\\

Stemmer\\

Tokenizer\\
Twitter data, and user-generated web content in general, provides unique challenges to an automatic tokenizing system due to the relatively small number of features present and high level of noise in the feature set [citation and more details]. Features on twitter include excessive and irregular punctuation, frequent misspellings, emoticons, URLs, emoticons, and hashtags, to name a few.

@@

Miscenalaneous\\

Unescape HTML\\

N-grams\\

Link2Title\\

Stopword Removal\\

\section{Results}

Discussion of accuracy etc

\section{Applications and Future Work}

@@

@@

\section{ Acknowledgments}

\begin{thebibliography}{99}
@article{potts2011,
    author    = "Christopher Potts",
    title     = "Sentiment Symposium Tutorial: Tokenizing",
    journal   = "",
    year      = "2011"
}
@inproceedings{Pak10,
  added-at = {2010-06-01T10:23:44.000+0200},
  address = {Valletta, Malta},
  author = {Pak, Alexander and Paroubek, Patrick},
  biburl = {http://www.bibsonomy.org/bibtex/25656c3bb1adf00c58a85e3204096961c/frederik},
  booktitle = {Proceedings of the Seventh conference on International Language Resources and Evaluation (LREC'10)},
  date = {19-21},
  interhash = {ac930b0459a3c8a2fc2d74c52a475026},
  intrahash = {5656c3bb1adf00c58a85e3204096961c},
  isbn = {2-9517408-6-7},
  keywords = {imported},
  language = {english},
  month = may,
  publisher = {European Language Resources Association (ELRA)},
  timestamp = {2010-06-01T10:23:44.000+0200},
  title = {Twitter as a Corpus for Sentiment Analysis and Opinion Mining},
  url = {http://www.lrec-conf.org/proceedings/lrec2010/pdf/385_Paper.pdf},
  year = 2010
}
\end{thebibliography}

>>>>

\title{Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets}
\author{
Curtis Ullerich, Daniel Stiner, Brandon Maxwell\\
Department of Computer Science and Engineering\\
Iowa State University
Ames, Iowa, USA\\
\{curtisu,stiner,bmaxwell\}@iastate.edu\\
}
\maketitle
\begin{abstract}
\begin{quote}
Twitter is a popular source for data mining due to its massive scale and inclusivity of current trends. As businesses seek to discover public opinion about their business or products, Twitter is one source for mining this information. Tweets present interesting classification challenges due to their irregular formatting and relatively small number of features. Tweet selection by keyword filtering often suffers from an high accuracy and low recall, or low accuracy and high recall. Using machine learning for text classification can improve this. We aim to survey a variety of text processing steps, focusing on the effects of tokenization, to increase accuracy of classification using a Naive Bayes classifier. We use a broad set of keywords with the goal of collecting a superset of the Tweets about a particular company (Apple, in this case) and train a binary classifier to filter out false positives.
\end{quote}
\end{abstract}

\section{Introduction}
Tweets are used as a source of data mining for applications ranging from predicting future stock price changes based on sentiment \cite{Ruiz:2012:CFT:2124295.2124358}, ... [ADD MORE]
In many cases, the problem of determining which tweets are relevant for a particular application is prerequisite to meaningful results. In some domains, filtering tweets based solely on keywords may be appropriate, but disambiguation remains an interest natural language processing problem in many cases. We present a system to disambiguate tweets about the company Apple.

\section{Related Work}
DUALIST is a system that extends Mallet by user-interactive feedback during training[Settles]. Included in their implementation was a TwitterPipe that performed some simple feature extraction. 
Part of the Sentiment Analysis Symposium focuses on tokenization of tweets for sentiment analysis, demonstrating that it provides consistent improvement over whitespace-based tokenizing \cite{potts2011}. A good deal of research has been done into sentiment classification of tweets, which requires very similar text processing and tokenization \cite{Pak10}.

\section{Method}
\subsection{Corpus creation}\\
By accessing the Twitter "Firehose" stream through their public API, we harvested 115,000 tweets using broad keyword filters with the goal of collecting a superset of relevant tweets. This provides us a smaller set of features in the overall corpus. The keywords used for the company of choice during this round of testing, Apple, were collected manually from information on the company's website and the Wikipedia page about Apple:\\
...\\
\\
This set of keywords clearly allows a large number of false positives:\\
nice fish and chip.. good pie apple :-) http://t.co/KVDN5r5l \\
20 chicken nuggets, chips, big mac, and a vanilla milkshake LIFE IS GOOD \\
1 more day whoop plus safari ride yay \\
shley staples an apple.  "AHHHH!!!!  Apple juice in my eye!" \\
I ain't seen uncle Mac in a while\\
\\
We define a tweet to be "about" Apple if the tweet mentions the company (via direct reference or stock symbol), one of its products, or a service it offers. We make the assumption that spam will be filtered from the testing set. This could be done via an existing Naive Bayes classifier, but we simply did not include instances of spam in our corpus during the manual labeling process. We collected our tweets over the course of twelve hours. On Twitter, a single message posted is often reposted, or "re-tweeted," many more times in a very short timespan. Because of the small timespan of our data collection, a large number of retweets (effectively duplicates) appeared in the dataset. We filtered these duplicates out of the dataset as well to prevent over-fitting in our model. After filtering, our original set of 115,000 tweets was reduced to 65,000. A representative sample of non-spam, non-duplicate tweets has composition of 60\% topical instances and 40\% false positives [need to analyze this for sure].
A computer-posted spam tweet: I've collected 10,650 gold coins! http://t.co/H0V0O6yP #ipad, #ipadgames, #gameinsight\\
A retweet, signified by "RT": RT @PHILerNotebook: Trying to fix my Mac. This gray loading bar takes 10 minutes to start up. Grr. Shall take it to an _Apple_ _store_ soon! \\
\\
We hand labeled 2000 tweets for use in training and testing, including only English-language tweets in our training set. \\
Data sets and scripts used for processing text data are available on the project website.

\subsection{Preprocessing Pipeline}

We do all machine learning with the Mallet[2] library developed by the University of Massachusetts. Mallet includes abstractions for a data processing pipeline, during which we use both existing and custom processing "Pipes" to transform the data, beginning in our case with raw tweets text and ending in Mallet's internal FeatureVector format. We built several Pipes to analyze the effect of different filters and processes on the accuracy of classification:

\subsubsection{Feature-space Enrichment}
Stopword removal:
N-grams:
Fix emoticons:
HTML corrections: 
URL replacement:
Spell checking:

\subsubsection{Disambiguation}
Lowercasing:
Stemming:

\subsubsection{Tokenization}
Twitter data, and user-generated web content in general, provides unique challenges to an automatic tokenizing system due to the relatively small number of features present and high level of noise in the feature set [citation and more details]. Features on twitter include excessive and irregular punctuation, frequent misspellings, emoticons, URLs, emoticons, and hashtags, to name a few.

@@

\section{Experimental Setup}
...constant-size test set with sliced training
...comparison to n-fold validation

\section{Results}
...baseline using priors
  baseline using whitespace splitting in naive bayes.
  select interesting improvements.
  graphs of both validation types
  most useful features for classification

\section{Conclusion and Future Work}

@@

[These guys] demonstrate a system of using background knowledge and relatedness, in which current information in the Twitter stream is used to more accurately estimate the priors for a Naive Bayes classification.

@@

\section{ Acknowledgments}
Thank you to Dr. Jin Tian and our TA Ru He for their guidance during this research project. We also extend our thanks to our friends for their help in manual labeling of tweets.

\section{References}
\begin{thebibliography}{99}
@article{potts2011,
    author    = "Christopher Potts",
    title     = "Sentiment Symposium Tutorial: Tokenizing",
    journal   = "",
    year      = "2011"
}
@inproceedings{Pak10,
  added-at = {2010-06-01T10:23:44.000+0200},
  address = {Valletta, Malta},
  author = {Pak, Alexander and Paroubek, Patrick},
  biburl = {http://www.bibsonomy.org/bibtex/25656c3bb1adf00c58a85e3204096961c/frederik},
  booktitle = {Proceedings of the Seventh conference on International Language Resources and Evaluation (LREC'10)},
  date = {19-21},
  interhash = {ac930b0459a3c8a2fc2d74c52a475026},
  intrahash = {5656c3bb1adf00c58a85e3204096961c},
  isbn = {2-9517408-6-7},
  keywords = {imported},
  language = {english},
  month = may,
  publisher = {European Language Resources Association (ELRA)},
  timestamp = {2010-06-01T10:23:44.000+0200},
  title = {Twitter as a Corpus for Sentiment Analysis and Opinion Mining},
  url = {http://www.lrec-conf.org/proceedings/lrec2010/pdf/385_Paper.pdf},
  year = 2010
}
@inproceedings{Ruiz:2012:CFT:2124295.2124358,
 author = {Ruiz, Eduardo J. and Hristidis, Vagelis and Castillo, Carlos and Gionis, Aristides and Jaimes, Alejandro},
 title = {Correlating financial time series with micro-blogging activity},
 booktitle = {Proceedings of the fifth ACM international conference on Web search and data mining},
 series = {WSDM '12},
 year = {2012},
 isbn = {978-1-4503-0747-5},
 location = {Seattle, Washington, USA},
 pages = {513--522},
 numpages = {10},
 url = {http://doi.acm.org/10.1145/2124295.2124358},
 doi = {10.1145/2124295.2124358},
 acmid = {2124358},
 publisher = {ACM},
 address = {New York, NY, USA},
 keywords = {financial time series, micro-blogging, social networks},
} 
\end{thebibliography}
